Evaluating the Performance and Energy Efficiency of the COSMO-ART model system

### 2014-04-25 paper_enahpc14/Introduction.tex
inserts related work

Hence to  improve our understanding of  aerosol-cloud interactions and
reduce  these  uncertainties,  the  research  community  is  making  a
concerted international  effort to represent  the underlying physical,
chemical  and aerosol dynamical  processes through  numerical chemical
transport  models (CTMs)  such  as ART  (Aerosols  and Reactive  Trace
gases),   developed   at  the   Karlsruhe   Institute  of   Technology
(KIT)  \citep{Vogel-2009,  Bangert-2011,  Knote-2013}.  This  regional
scale model  system is coupled  with the operational  weather forecast
model COSMO \citep{Baldauf-2011}, jointly developed by a consortium of
European weather  centers and used in the  climate version (COSMO-CLM)
by a wide research community.

The extended atmospheric model  COSMO-ART is computationally much more
demanding than  COSMO since a  large number of additional  tracers and
processes  have  to  be  considered.   Thus  the  model  is  currently
severely-limited in  terms of applicability and expensive  in terms of
energy  consumption.   Although  COSMO  has recently  been  ported  to
GPUs  \citep{Gysi-2014, Lapillonne-2014} within  the framework  of the
High    Performance   and    High   Productivity    Computing   (HP2C)
Initiative \citep{HP2C},  to optimise it for  computational and energy
efficiency, significant investments in  ART are still required to take
it to a similar level.  The efficiency of COSMO-ART is being addressed
in the  EU Exa2Green project \citep{EXA2GREEN} to  deliver a prototype
code, which  provides an energy efficiency  of at least  five times of
the baseline value.  Such  an implementation would allow the community
to investigate critical questions at higher resolution and over longer
periods, at reduced cost to the environment.

This  work is  organized as  follows: in  Sec.~\ref{sec:1} we  give an
overview  of  related  work,   then  in  Sec.~\ref{sec:2}  we  briefly
introduce  COSMO-ART and specify  its technical  setup related  to the
investigated   performance   and   energy  evaluation   methods.    In
Sec.~\ref{sec:3}   we  describe   HPC  platforms,   power  measurement
equipments   and  software  environment   employed  to   conduct  this
benchmarking study.  Sec.~\ref{sec:4}  presents performance and energy
requirements  of the  baseline on  these architectures  and highlights
areas where improvements will be necessary for the subsequent baseline
refactoring.   Finally,  we  conclude  with  practical  guidelines  to
identifying energy saving opportunities  and give an outlook to future
research in Sec.~\ref{concl}.

>>>>

Hence to  improve our understanding of  aerosol-cloud interactions and
reduce  these  uncertainties,  the  research  community  is  making  a
concerted international  effort to represent  the underlying physical,
chemical  and aerosol dynamical  processes through  numerical chemical
transport  models (CTMs)  such  as ART  (Aerosols  and Reactive  Trace
gases),   developed   at  the   Karlsruhe   Institute  of   Technology
(KIT)  \citep{Vogel-2009,  Bangert-2011,  Knote-2013}.  This  regional
scale model  system is coupled  with the operational  weather forecast
model  \textsc{Cosmo}  \citep{Baldauf-2011},  jointly developed  by  a
consortium of European weather centers and used in the climate version
by a wide research community.

The extended  atmospheric model \textsc{Cosmo-art}  is computationally
much  more  demanding than  \textsc{Cosmo}  since  a  large number  of
additional  tracers and  processes have  to be  considered.   Thus the
model  is currently  severely-limited  in terms  of applicability  and
expensive in terms of energy consumption.  Although \textsc{Cosmo} has
recently been ported to GPUs \citep{Gysi-2014, Lapillonne-2014} within
the framework of the  High Performance and High Productivity Computing
(HP2C) Initiative  \citep{HP2C}, to optimise it  for computational and
energy efficiency,  significant investments in ART  are still required
to take it  to a similar level.  The  efficiency of \textsc{Cosmo-art}
is being  addressed in the  EU Exa2Green project  \citep{EXA2GREEN} to
deliver a  prototype code, which  provides an energy efficiency  of at
least five times of the  baseline value.  Such an implementation would
allow  the  community  to  investigate critical  questions  at  higher
resolution  and   over  longer  periods,   at  reduced  cost   to  the
environment.

This  work is  organized as  follows: in  Sec.~\ref{sec:1} we  give an
overview  of  related  work,   then  in  Sec.~\ref{sec:2}  we  briefly
introduce \textsc{Cosmo-art}  and specify its  technical setup related
to  the investigated  performance and  energy evaluation  methods.  In
Sec.~\ref{sec:3}   we  describe   HPC  platforms,   power  measurement
equipments   and  software  environment   employed  to   conduct  this
benchmarking study.  Sec.~\ref{sec:4}  presents performance and energy
requirements  of the  baseline on  these architectures  and highlights
areas where improvements will be necessary for the subsequent baseline
refactoring.   Finally,  we  conclude  with  practical  guidelines  to
identifying energy saving opportunities  and give an outlook to future
research in Sec.~\ref{concl}.

### 2014-04-25 paper_enahpc14/Sec1:Related.tex
inserts related work

\begin{itemize}
\item Application centric energy-efficiency study of distributed
  multi-core and hybrid {CPU-GPU} systems \citep{Cumming-2014}
\item Stella: A domain-specific language and tool for structured grid
  methods \citep{Gysi-2014}
\item Using compiler directives to port large scientific applications
  to GPUs: An example from atmospherci science \citep{Lapillonne-2014}
\item An analysis of energy-optimized lattice-Boltzmann CFD
  simulations from the chip to the highly parallel level
  \citep{Wittmann-2013}
\item Energy efficiency vs. performance of the numerical solution of
  PDEs: An application study on a low-power ARM-based cluster
  \citep{Goddeke-2013}
\item Analysis of Computing and Energy Performance of Multicore, NUMA,
  and Manycore Platforms for an Irregular Application
  \citep{Castro-2013}
\item Exploring performance and power properties of modern multi-core
  chips via simple machine models \citep{Hager-2013}
\end{itemize}

>>>>

While the  TOP500 list was  introduced over 20  years ago to  rank the
performance of HPC systems  worldwide, it's quite recently that energy
efficiency  become  a  critical  constraint  in the  way  to  exascale
computing.   Since 2007,  the Green500  power  measurement methodology
encourages the design,  procurement and management of energy-efficient
infrastuctures to contain performance in an affordable and competitive
power  envelope.   However,  the state-of-the-art  research  assessing
performance  and energy  efficiency  of applications  is still  pretty
scarce.

\emph{Padoin et al.}   \citep{Padoin-2013} investigate performance and
power  consumption  of  an  agroforestry  application  and  show  that
changing workload  can drastically improve energy efficiency  of CPU +
GPU heterogeneous architectures.

\emph{Ou  Pang et  al.}   \citep{Ou-2012} compare  ARM  and Intel  x86
workstations  clusters  and   conclude  that  ARM-based  clusters  are
advantageous  with   lightweight  applications  in   terms  of  energy
efficiency.

\emph{G\"oddeke et al.}  \citep{Goddeke-2013} evaluate weak and strong
scalability of PDE solvers on a cluster of 96 ARM dual-core processors
and demonstrate  that the ARM-based  cluster can be more  efficient in
terms of energy-to-solution compared to x86-based cluster.

\emph{Wittmann  et  al.}   \citep{Wittmann-2013}  perform  a  thorough
analysis of  a lattice-Boltzmann method based CFD  simulation on Intel
Sandy   Bridge  processors   and  show   extrapolated  results   on  a
petascale-class machine.

\emph{Cumming  et  al.}   \citep{Cumming-2014}  present a  simple  and
practical  methodology  looking at  energy  minimisation  that can  be
applied to various applications.

### 2014-04-25 paper_enahpc14/Sec3:Environment-setup.tex
inserts related work

We have  chosen a state-of-the-art Intel's  third-generation Core (aka
Ivy  Bridge) processing platform  at CSCS  (called ``Monch'')  for our
power  measurements,  which  is  slated  to stay  in  service  without
hardware  upgrade  for the  duration  of  the  Exa2Green project.   In
principle at least,  this architecture could be recreated  or found in
an identical configuration beyond  the lifetime of the project.  Since
the baseline benchmark can  be reproduced within an expected variance,
and  that the baseline  run configuration  can be  used in  all future
versions  of the  code, a  fair comparison  will be  made  between the
baseline and  the milestone versions  of COSMO-ART.  In this  study, a
complementary  energy-to-solution benchmarking  comparison  is carried
out   on  the   ``Pilatus''  cluster   at  CSCS,   based   on  Intel's
previous-generation  Sandy  Bridge  processors,  conventional  in  HPC
systems and known to be  more power consuming.  Finally a full tracing
experiment is  conducted on the  ``Tintorrum'' cluster at the  Jaume I
University (UJI) using an  integrated framework, to capture an overall
power profile at a much finer resolution and to get more insights into
the MPI blocking and polling influences on power savings.

\subsubsection{CSCS - Monch}
Monch is  a 10 rack  NEC-provided and dual-socket Intel  Ivy Bridge-EP
based  cluster, utilised  by  scientists  who are  part  of the  Swiss
Platform     for      Advanced     Scientific     Computing     (PASC,
\url{http://www.pasc-ch.org/}). It is composed of 312 standard compute
nodes, 24 large-memory compute nodes and 24 huge-memory compute nodes.
Each standard  compute node comprises  two Intel Ivy  Bridge Efficient
Performance (EP)  E5-2660 v2 ten-core processors operating  at 2.2 GHz
base  clock speed,  themselves connected  by a  high  speed InfiniBand
network based on Mellanox SX6036  managed FDR switches, with a 56 Gb/s
speed.  Each core  has 32 KB instruction and 32 KB  data L1 caches and
256  KB of L2  cache. All  10 cores  share a  25 MB  L3 cache  and the
platform has  32GB of  DDR3 1600 MHz  RAM. For  our energy-to-solution
benchmark, a  full rack  of Monch constituted  of 52  standard compute
nodes (monchc[029-080]) was considered.

\subsubsection{CSCS - Pilatus} 
Pilatus  is  a dual-socket  eight-core  Intel  Sandy  Bridge EP  based
cluster used as Piz Daint pre-post processing cluster.  It is composed
of 42 compute  nodes and has 2 high-speed  interconnects based on FDR:
the  first is  dedicated to  the  MPI traffic  and the  second to  the
storage high  speed traffic.   The 2 login  nodes and the  42 computes
nodes  consists in  11  twin-pair Intel  E5-Series  DALCO r2264i4t  2U
scalable compute modules.  Each  module contains 4 compute nodes based
on two Intel  Xeon E5-2670 processors operating at  2.6 GHz base clock
speed, themselves  connected by a high speed  InfiniBand network based
on Mellanox  SX6036 managed FDR switches,  with a 56  Gb/s speed. Each
core has 32 KB  instruction and 32 KB data L1 caches  and 256 KB of L2
cache.  All  the 8 core share  a 20 MB  L3 cache and the  platform has
64GB of  DDR3 1600 MHz  RAM.  For our energy-to-solution  benchmark, a
full  rack  of  Pilatus  constituted  of  42  standard  compute  nodes
(pilatus[03-44]) was considered.

\subsubsection{UJI - Tintorrum} 
Tintorrum  is a heterogeneous  cluster composed  of 28  compute nodes.
For our experiments only a subset of these nodes were considered. This
set is  composed of 16  nodes, each of  which includes two  Intel Xeon
E5645 hexa-core  processors at 2.40  GHz connected via  Infiniband QDR
(Mellanox MTS3600 switch).  Each core  has 32 KB instruction and 32 KB
data L1 caches and 256 KB L2 cache. The 6 cores share a 12 MB L3 cache
and the platform has 288 GB of DDR3 1333 MHz.

The COSMO-ART baseline  is a pure MPI based  Fortran~90 code currently
running on distributed multi-core systems only.  The software stack on
both CSCS  platforms was controlled using the  modules framework which
gives an easy and flexible mechanism  to access to all of the provided
compilers, tools  and applications.  For our  initial benchmarking, we
opted for the GNU compiler  (gcc/4.8.1 on Monch, gcc/4.8.2 on Pilatus)
using the -O3  compiler flag in favor of  the intel compiler (14.0.1),
delivering inferior performance.  In addition, we installed the MPICH2
implementation of MPI (mvapich2/1.9) as well as the commonly used HDF5
(1.8.12) and NetCDF (4.3.1)  libraries for the management of extremely
large  and  complex data  collections.   All  computes  nodes have  an
operating      system      based      on      GNU/Linux      featuring
``2.6.32-358.11.1.el6.x86\_64''      kernel      in     Monch      and
``3.0.101-0.15-default'' kernel in Pilatus.

\subsubsection{Run configuration}
A snapshot of the code, which includes, at least conceptually, all the
information needed  to reproduce the  energy-to-solution benchmarks of
COSMO-ART, was produced and run on a 1040 cores using 20 MPI tasks per
node on  Monch and  on a  1344 cores using  16 MPI  tasks per  node on
Pilatus.   The  calculated  region  was mapped  to  the  participating
processors using  a 2D-partitioning strategy.   The distribution along
the $x$  and $y$ coordinates  was defined by setting:  $nprocx=40$ and
$nprocy=26$ for  Monch and $nprocx=28$ and $nprocy=24$  as $nprocx$ is
usually  kept  bigger than  $nprocy$.   Besides,  as  this version  of
COSMO-ART  doesn't  make  use  of  the traditional  GRIB  library,  we
specified $nprocio=0$ for GRIB  I/O.  Hyperthreading is not considered
in this study as previous attempts  of its use revealed that it always
led to higher energy-to-solution.\\

@@

\subsubsection{CSCS - E3METER metering products}
Supercomputer  clusters considered  for our  experiments at  the Swiss
National Supercomputing Center (CSCS)  of ETH Zurich are equipped with
E3METER Intelligent  Power Strips (IPS)  and Monitors (IPM)  which are
high   quality   electricity  meters   released   by  Riedo   Networks
(\url{http://riedonetworks.com/}),  that  enable  to monitor  and  log
power  consumption of  the  IT infrastructure  as  well as  constantly
analyze  line  voltage,  current,  power-factor,  frequency  with  1\%
accuracy.   Using reliable  narrowband  powerline communication  (PLC)
technology,  all metering  and power  quality data  from each  IPS are
centrally collected by the E3METER Data Concentrator, via the existing
power cables thus  avoiding the need for extra  cabling.  This data is
made  available  via SNMP,  HTTP,  TELNET  through  the built-in  Fast
Ethernet  port.   Time  synchronisation  is guaranteed  by  using  NTP
servers.   Measured data  is accessed  through the  open  source Cacti
software including  the E3METER  Cacti Plugin to  scan the  entire PLC
network and monitor  in real-time the power usage  of individual rack,
recorded in 5 minutes interval periods.

\subsubsection{UHAM - UJI}
To assess the  performance and the energy efficiency  of COSMO-ART, we
employ   a    version   of   the    integrated   framework   presented
in~\cite{energy13} that works in  combination with Extrae and Paraver,
which are profiling/tracing and visualization tools, respectively.

%The left part of the \vref{fig:Lustre} offers a graphical representation of
%the Lustre architecture; the right depicts the tracing and profiling framework.
To use our  approach, COSMO-ART is compiled using  the Extrae compiler
wrappers,  which  automatically instrument  the  Fortran  code of  the
model. Next, COSMO-ART  is run on the nodes,  thus dissipating certain
amount  of power.   These  nodes are  connected  to power  measurement
devices that account for the dissipated power/consumed energy and send
the power  data to the  tracing server.  The client,  meanwhile, sends
start/stop  primitives  in  order  to  gather  captured  data  by  the
watt-meters onto the  tracing server, where an instance  of the \pmlib
server is running.

Once COSMO-ART run is finished, a file containing the power profile is
created using the power data  received from the tracing server and the
instrumentation post-processing generates the performance trace files.
All  these files  are  combined then  in  Paraver which  allows us  to
visualise the performance trace and the power profile of COSMO-ART all
together.

%Once COSMO-ART run is finished, the VampirTrace \pmlib plugin receives
%the  power   data  from  the  tracing   server.   The  instrumentation
%post-processing generates  the performance trace files  and the \pmlib
%plugin inserts the power data into them.

@@

>>>>

Benchmarks presented  in this work  were performed on  three different
platforms  and can  be reproduced  within an  expected  variance, thus
allowing a  fair comparison between the baseline  and future milestone
versions of \textsc{Cosmo-art}.

\begin{itemize}
\item \textsc{Monch (CSCS)} is a 10 rack NEC-provided cluster composed
  of 312 compute  nodes and utilised by members  of the Swiss Platform
  for Advanced Scientific  Computing (PASC) project \citep{PASC}. Each
  node comprises  two Intel Ivy Bridge EP  E5-2660 ten-core processors
  operating at  2.2 GHz connected  via InfiniBand Mellanox  SX6036 and
  FDR switches (56 Gb/s speed), and is equipped with 32GB of DDR3 1600
  MHz  RAM.   For our  experiments,  a  full  rack constituted  of  52
  standard compute  nodes was used.  \textsc{Monch} is  slated to stay
  in  service  without  hardware  upgrade  for  the  duration  of  the
  Exa2Green  project,  allowing   an  identical  configuration  to  be
  recreated for future assessments of the baseline.\\

\item \textsc{Pilatus (CSCS)} is  a dual-socket eight-core Intel Sandy
  Bridge EP based cluster composed of 42 compute nodes and used as Piz
  Daint pre-post  processing cluster at  CSCS.  The 2 login  nodes and
  the 42 computes nodes consists in 11 twin-pair Intel E5-Series DALCO
  r2264i4t 2U scalable compute modules. Each module contains 4 compute
  nodes based  on two Intel  Xeon E5-2670 processors operating  at 2.6
  GHz connected  via InfiniBand Mellanox  SX6036 and FDR  switches (56
  Gb/s speed), and is equipped with 64GB of DDR3 1600 MHz RAM. For our
  experiments, a  full rack constituted  of 42 standard  compute nodes
  was  used. \textsc{Pilatus}  is based  on  Intel's second-generation
  processors, conventional in  HPC systems and known to  be more power
  consuming than Ivy Bridge successor.\\

\item \textsc{Tintorrum (UJI)} is  a heterogeneous cluster composed of
  28 compute nodes.  For our  experiments only a subset of these nodes
  were considered.   This set is composed  of 16 nodes,  each of which
  includes  two Intel  Xeon  E5645 hexa-core  processors  at 2.40  GHz
  connected via  Infiniband QDR (Mellanox MTS3600  switch).  Each core
  has  32 KB  instruction  and 32  KB data  L1  caches and  256 KB  L2
  cache. The 6 cores  share a 12 MB L3 cache and  the platform has 288
  GB  of DDR3 1333  MHz.  A  full tracing  experiment is  conducted on
  \textsc{Tintorrum}, to  capture an overall  power profile at  a much
  finer resolution, and to get more insights into the MPI blocking and
  polling influences on power savings.
\end{itemize}

@@

\subsubsection{CSCS - E3METER metering products}

Supercomputer  clusters considered  for our  experiments at  the Swiss
National Supercomputing Center (CSCS)  of ETH Zurich are equipped with
E3METER Intelligent  Power Strips (IPS)  and Monitors (IPM)  which are
high   quality   electricity  meters   released   by  Riedo   Networks
\citep{Riedonetworks},   that  enable   to  monitor   and   log  power
consumption  of the IT  infrastructure as  well as  constantly analyze
line  voltage,  current, power-factor,  frequency  with 1\%  accuracy.
Using  reliable narrowband  powerline communication  (PLC) technology,
all  metering and  power  quality  data from  each  IPS are  centrally
collected  by the E3METER  Data Concentrator,  via the  existing power
cables thus  avoiding the need for  extra cabling.  This  data is made
available via  SNMP, HTTP, TELNET  through the built-in  Fast Ethernet
port.   Time  synchronisation  is  guaranteed by  using  NTP  servers.
Measured  data is  accessed  through the  open  source Cacti  software
including the E3METER Cacti Plugin  to scan the entire PLC network and
monitor in real-time the power usage of individual rack, recorded in 5
minutes interval periods.

\subsubsection{UHAM - UJI}

For our experiments at Jaume I University (UJI), we employed a version
of the integrated framework presented in~\cite{energy13} that works in
combination with  Extrae and Paraver, which  are profiling/tracing and
visualization tools, respectively.

%The left part of the \vref{fig:Lustre} offers a graphical representation of
%the Lustre architecture; the right depicts the tracing and profiling framework.
\textsc{Cosmo-art}  is compiled  using the  Extrae  compiler wrappers,
which automatically  instrument the Fortran code of  the model.  Next,
\textsc{Cosmo-art}  is  run on  the  nodes,  thus dissipating  certain
amount  of power.   These  nodes are  connected  to power  measurement
devices that account for the dissipated power/consumed energy and send
the power  data to the  tracing server.  The client,  meanwhile, sends
start/stop  primitives  in  order  to  gather  captured  data  by  the
watt-meters onto the  tracing server, where an instance  of the \pmlib
server is running.

Once  the run  is finished,  a file  containing the  power  profile is
created using the power data  received from the tracing server and the
instrumentation post-processing generates the performance trace files.
All  these files  are  combined then  in  Paraver which  allows us  to
visualise   the   performance  trace   and   the   power  profile   of
\textsc{Cosmo-art} all together.

%Once \textsc{Cosmo-art} run is finished, the VampirTrace \pmlib plugin receives
%the  power   data  from  the  tracing   server.   The  instrumentation
%post-processing generates  the performance trace files  and the \pmlib
%plugin inserts the power data into them.

@@

The \textsc{Cosmo-art}  baseline is a  pure MPI based  Fortran~90 code
currently  running  on   distributed  multi-core  systems  only.   The
software stack on both CSCS platforms was controlled using the modules
framework which gives an easy  and flexible mechanism to access to all
of the  provided compilers, tools  and applications.  For  our initial
benchmarking,   we  opted   for   the  GNU   compiler  (gcc/4.8.1   on
\textsc{Monch}, gcc/4.8.2 on  \textsc{Pilatus}) using the -O3 compiler
flag  in favor  of the  intel compiler  (14.0.1),  delivering inferior
performance.  In  addition, we installed the  MPICH2 implementation of
MPI  (mvapich2/1.9) as  well as  the commonly  used HDF5  (1.8.12) and
NetCDF  (4.3.1) libraries for  the management  of extremely  large and
complex data collections.  All computes nodes have an operating system
based on GNU/Linux featuring ``2.6.32-358.11.1.el6.x86\_64'' kernel in
\textsc{Monch}     and      ``3.0.101-0.15-default''     kernel     in
\textsc{Pilatus}.

A snapshot of the code,  which includes at least conceptually, all the
information needed  to reproduce the  energy-to-solution benchmarks of
\textsc{Cosmo-art}, was produced and run  on a 1040 cores using 20~MPI
tasks  per node on  \textsc{Monch} and  on a  1344 cores  using 16~MPI
tasks per node on  \textsc{Pilatus}.  The calculated region was mapped
to the participating processors using a 2D-partitioning strategy.  The
distribution along the $x$ and $y$ coordinates was defined by setting:
$nprocx=40$  and $nprocy=26$  for \textsc{Monch}  and  $nprocx=28$ and
$nprocy=24$  as  $nprocx$  is   usually  kept  bigger  than  $nprocy$.
Besides,  as this  version doesn't  make use  of the  traditional GRIB
library, we  specified $nprocio=0$  for GRIB I/O.   Hyperthreading was
not considered in this study  as it previously revealed that it always
led to higher energy-to-solution.

### 2014-04-25 paper_enahpc14/paper.tex
inserts related work

\begin{abstract}
  In this  paper we investigate  the energy footprint  and performance
  profiling of COSMO-ART  on various HPC platforms.  This  model is an
  extension of  the operational weather  forecast model of  the German
  weather  service   (DWD),  developed  for  the   evaluation  of  the
  interactions of reactive gases  and aerosol particles with the state
  of atmosphere at the  regional scale.  Different measurement devices
  and energy-aware techniques are  described to evaluate both time and
  energy  to  solution  of  the  considered application  and  to  gain
  detailed  insights  into power  and  performance requirements.   Our
  motivation  is to  improve  corresponding code  sections to  sustain
  performance  while minimising energy-to-solution.   This preliminary
  work  sets the  basis for  subsequent studies  to  tackle challenges
  related  to  energy  efficient  high performance  computing  in  the
  framework of the Exa2Green project \citep{EXA2GREEN}.

\keywords{High performance computing  \and Energy-aware computing \and
  Green computing  \and Numerical weather  prediction \and Atmospheric
  chemistry  \and  Aerosols  modelling  \and  Profiling  methods  \and
  Benchmark    analysis   \and    COSMO-ART    coupled   model    \and
  Time-to-solution \and Energy-to-solution \and Xeon processor}
% \PACS{PACS code1 \and PACS code2 \and more}
% \subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}

@@

\section{The COSMO-ART model system}
\label{sec:2}
\input{Sec2:COSMO-ART.tex}

>>>>

\begin{abstract}
  In this  paper we investigate  the energy footprint  and performance
  profiling  of  \textsc{COSMO-ART} on  various  HPC platforms.   This
  model is an  extension of the operational weather  forecast model of
  the German  weather service (DWD),  developed for the  evaluation of
  the interactions  of reactive gases  and aerosol particles  with the
  state of  atmosphere at  the regional scale.   Different measurement
  devices and  energy-aware techniques are described  to evaluate both
  time and  energy to  solution of the  considered application  and to
  gain detailed insights into power and performance requirements.  Our
  motivation  is to  improve  corresponding code  sections to  sustain
  performance  while minimising energy-to-solution.   This preliminary
  work  sets the  basis for  subsequent studies  to  tackle challenges
  related  to  energy  efficient  high performance  computing  in  the
  framework of the Exa2Green project \citep{EXA2GREEN}.

\keywords{High performance computing  \and Energy-aware computing \and
  Green computing  \and Numerical weather  prediction \and Atmospheric
  chemistry  \and  Aerosols  modelling  \and  Profiling  methods  \and
  Benchmark  analysis  \and   \textsc{COSMO-ART}  coupled  model  \and
  Time-to-solution \and Energy-to-solution \and Xeon processor}
% \PACS{PACS code1 \and PACS code2 \and more}
% \subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}

@@

\section{The \textsc{COSMO-ART} model system}
\label{sec:2}
\input{Sec2:COSMO-ART.tex}
